My wonderful paper

### 2024-10-31 _manuscript/_tex/index.tex
second version of the introduction

\begin{document}
\maketitle
\begin{abstract}
This is the abstract
\end{abstract}

@@

{[}background - In a data analysis workflow{]}

In a data analysis workflow -\textgreater{} going back to diagnose the
descripency between the expectation and the result is often
time-consuming and challenging.

For consumers of data analysis, it is often difficult to evaluate the
analysis outcome. Not always have the script to run, even difficult when
different expectations are held.

We propose here a concept called analysis plan to make the expectation
of the analysis process explicit. It can help to guide the analysis
process, to communicate and compare the analysis process to others, and
to evaluate the analysis process. (think more closer here how our
approach help to solve the problem)

For expert analysts, the expectation of the analysis process is often
implicit. They may have a mental model of the analysis process and the
expected results. However, for non-expert analysts, the expectation of
the analysis process may not be clear. They may not know what to expect
from the analysis process, or what to look for when an analysis process
fails the expectation. An analysis plan can help to make the expectation
of the analysis process explicit. It can help to guide the analysis
process, to communicate and compare the analysis process to others, and
to evaluate the analysis process.

In this paper, a concept called analysis plan is proposed to describe
the logical structure of a data analysis. An analysis plan is a set of
analysis steps plus their expected outcomes. It is a formal
representation of the analysis process and can be used to guide the
analysis process, to communicate and compare the analysis process to
others, and to evaluate the analysis process. The concept of analysis
plan is illustrated with examples. The implications of the concept for
data analysis practice is discussed.

The analysis plan described in this paper should be differentiated from
the pre-specifies analysis plan document often used in biostatistics to
specifies the hypothesis, data collection mechanism, statistical
procedures etc of randomized experiments.

@@

@@

TODO: more about analysis plan itself here

The whole point being how can be build trust in other's analysis. Some
provides scripts to run the analysis but that's not enough. We need to
know what to expect from the analysis and that is often not explicitly
specified.

>>>>

\begin{document}
\maketitle
\begin{abstract}
Unexpected results in data analysis can prompt researchers to examine
data quality and analysis steps. While some researchers can typically
diagnose issues effectively with a few checks, others may struggle to
identify appropriate checks to diagnose the problem. {[}an example of
check{]} These checks are often informal and difficult to trace and
discuss in publications, resulting in others questioning the
trustworthiness of the analysis. To address this, we formalize the
informal checks into an \emph{analysis plan} that encompasses the
analysis steps and (the unit tests): one test for whether the result
meets expectations and multiple tests for checking the analysis. We then
present a procedure to assess the quality of these unit tests based on
their accuracy and redundancy on simulated versions of the original
data. The accuracy is assessed using binary classification metrics,
\emph{i.e.}, precision and recall, while redundancy is calculated using
mutual information. This procedure can be used to conduct a sensitivity
analysis, compare different analysis plans, and to identify the optimal
cutoff point for the unit tests.
\end{abstract}

@@

In a data analysis, an experienced researcher can often quickly assess
whether a result meets their expectation, while others may be unsure of
what to anticipate from the analysis or how to recognize when results
diverge form expected norms. (These expectations are often based on
prior knowledge, domain expertise, or common sense and they can often be
framed into a Boolean question that can be answered by a unit test. For
example, in a linear regression model, one may expect the p-value of the
coefficient to be less than 0.05. On top of the expectation on the final
results, researchers may also have expectations on the intermediate
steps of the analysis process. For example, one may expect the data
doesn't contain outliers, etc. These expectations are often checked
early on to prevent an unexpected propagate through the analysis
process.)

This expectation-setting process is often implicit and rarely documented
or discussed in publications. (There are most publication requires open
source code for publication, however, it is no instruction or guidance
on checking \ldots). Yet, making them explicit is crucial for 1) helping
junior researchers interpret results and diagnose the unexpected, 2)
providing checkpoints that support independent replication or
application of methods to new data, 3) aligning assumptino across
researchers from different backgrounds.

In this paper, we formalize these informal checks into an \emph{analysis
plan} comprising the analysis steps along with the associated unit
tests. This formalization allows us to examine assumptions made about
the data during analysis and to compare different unit tests in
diagnosing unexpected outcomes. We then introduce a procedure to assess
the quality of these unit tests based on their accuracy and redundancy
across simulated version of the original data. The accuracy is assessed
using binary classification metrics -- precision and recall -- while
redundancy is measured using mutual information. We illustrate these
concepts with a simple step count example and a scenario of diagnosing
unexpected regression results when studying the effect of PM10 on
mortality.

@@

\begin{itemize}
\tightlist
\item
  \texttt{pointblank}, \texttt{assertthat}
\end{itemize}

@@

### 2024-10-31 index.qmd
second version of the introduction

---
title: "My wonderful paper"
abstract: "This is the abstract"
cite-method: natbib
bibliography: references.bib
documentclass: jds
format:
  pdf:
   keep-tex: true
   toc: true
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
editor_options: 
  chunk_output_type: console
---

@@

[background - In a data analysis workflow]

In a data analysis workflow -> going back to diagnose the descripency between the expectation and the result is often time-consuming and challenging.

For consumers of data analysis, it is often difficult to evaluate the analysis outcome. Not always have the script to run, even difficult when different expectations are held.

We propose here a concept called analysis plan to make the expectation of the analysis process explicit. It can help to guide the analysis process, to communicate and compare the analysis process to others, and to evaluate the analysis process. (think more closer here how our approach help to solve the problem)

For expert analysts, the expectation of the analysis process is often implicit. They may have a mental model of the analysis process and the expected results. However, for non-expert analysts, the expectation of the analysis process may not be clear. They may not know what to expect from the analysis process, or what to look for when an analysis process fails the expectation. An analysis plan can help to make the expectation of the analysis process explicit. It can help to guide the analysis process, to communicate and compare the analysis process to others, and to evaluate the analysis process.

In this paper, a concept called analysis plan is proposed to describe the logical structure of a data analysis. An analysis plan is a set of analysis steps plus their expected outcomes. It is a formal representation of the analysis process and can be used to guide the analysis process, to communicate and compare the analysis process to others, and to evaluate the analysis process. The concept of analysis plan is illustrated with examples. The implications of the concept for data analysis practice is discussed.

The analysis plan described in this paper should be differentiated from the pre-specifies analysis plan document often used in biostatistics to specifies the hypothesis, data collection mechanism, statistical procedures etc of randomized experiments.

@@

@@

TODO: more about analysis plan itself here

The whole point being how can be build trust in other's analysis. Some provides scripts to run the analysis but that's not enough. We need to know what to expect from the analysis and that is often not explicitly specified.

>>>>

---
title: "My wonderful paper"
abstract: "Unexpected results in data analysis can prompt researchers to examine data quality and analysis steps. While some researchers can typically diagnose issues effectively with a few checks, others may struggle to identify appropriate checks to diagnose the problem. [an example of check] These checks are often informal and difficult to trace and discuss in publications, resulting in others questioning the trustworthiness of the analysis. To address this, we formalize the informal checks into an *analysis plan* that encompasses the analysis steps and (the unit tests): one test for whether the result meets expectations and multiple tests for checking the analysis. We then present a procedure to assess the quality of these unit tests based on their accuracy and redundancy on simulated versions of the original data. The accuracy is assessed using binary classification metrics, *i.e.*, precision and recall, while redundancy is calculated using mutual information. This procedure can be used to conduct a sensitivity analysis, compare different analysis plans, and to identify the optimal cutoff point for the unit tests."
cite-method: natbib
bibliography: references.bib
documentclass: jds
format:
  pdf:
   keep-tex: true
   toc: true
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
editor_options: 
  chunk_output_type: console
---

@@

In a data analysis, an experienced researcher can often quickly assess whether a result meets their expectation, while others may be unsure of what to anticipate from the analysis or how to recognize when results diverge form expected norms. (These expectations are often based on prior knowledge, domain expertise, or common sense and they can often be framed into a Boolean question that can be answered by a unit test. For example, in a linear regression model, one may expect the p-value of the coefficient to be less than 0.05. On top of the expectation on the final results, researchers may also have expectations on the intermediate steps of the analysis process. For example, one may expect the data doesn't contain outliers, etc. These expectations are often checked early on to prevent an unexpected propagate through the analysis process.)

This expectation-setting process is often implicit and rarely documented or discussed in publications. (There are most publication requires open source code for publication, however, it is no instruction or guidance on checking ...).  Yet, making them explicit is crucial for 1) helping junior researchers interpret results and diagnose the unexpected, 2) providing checkpoints that support independent replication or application of methods to new data, 3) aligning assumptino across researchers from different backgrounds.

In this paper, we formalize these informal checks into an *analysis plan* comprising the analysis steps along with the associated unit tests. This formalization allows us to examine assumptions made about the data during analysis and to compare different unit tests in diagnosing unexpected outcomes. We then introduce a procedure to assess the quality of these unit tests based on their accuracy and redundancy across simulated version of the original data. The accuracy is assessed using binary classification metrics -- precision and recall -- while redundancy is measured using mutual information. We illustrate these concepts with a simple step count example and a scenario of diagnosing unexpected regression results when studying the effect of PM10 on mortality.

@@

- `pointblank`, `assertthat`

@@

<!-- The analysis plan described in this paper should be differentiated from the pre-specifies analysis plan document often used in biostatistics to specifies the hypothesis, data collection mechanism, statistical procedures etc of randomized experiments.  -->
